---
title: "ggcoef: Plot Model Coefficients with broom and ggplot2"
author: Joseph Larmarange
date: May 16, 2016
output:
  html_document:
    highlight: default
    toc: true
    theme: united
vignette: >
  %\VignetteIndexEntry{ggcoef}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# ggcoef

The purpose of this function is to quickly plot the coefficients of a model.

## Quick coefficients plot

To work automatically, this function requires the `broom` package. Simply call `ggcoef` with a model object. It could be the result of `lm`, `glm` or any other model covered by `broom` and its `tidy` method^[See http://www.rdocumentation.org/packages/broom.].

```{r}
library(GGally)
library(ggplot2)
reg <- lm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width, data = iris)
ggcoef(reg)
```

In the case of a logistic regression (or any other model for which coefficients are usually exponentiated), simply indicated `exponentiate = TRUE`. Note that a logarithmic scale will be used for the x-axis.

```{r}
d <- as.data.frame(Titanic)
log.reg <- glm(Survived ~ Sex + Age + Class, family = binomial, data = d, weights = d$Freq)
ggcoef(log.reg, exponentiate = TRUE)
```

## Customizing the plot

You can use `conf.int`, `vline` and `exclude_intercept` to display or not confidence intervals as error bars, a vertical line for `x = 0` (or `x = 1` if coeffcients are exponentiated) and the intercept.

```{r}
ggcoef(reg, vline = FALSE, conf.int = FALSE, exclude_intercept = TRUE)
```

See the help page of `ggcoef` for the full list of arguments that could be used to personalize how error bars and the vertical line are plotted.

```{r}
ggcoef(
  log.reg,
  exponentiate = TRUE,
  vline_color = "red",
  vline_linetype =  "solid",
  errorbar_color = "blue",
  errorbar_height = .25
)
```

Additional parameters will be passed to `geom_point`.

```{r}
ggcoef(log.reg, exponentiate = TRUE, color = "purple", size = 5, shape = 18)
```

Finally, you can also customize the aesthetic mapping of the points.

```{r}
ggcoef(log.reg, exponentiate = TRUE, mapping = aes(x = estimate, y = term, size = p.value)) +
  scale_size_continuous(trans = "reverse")
```

## Custom data frame

You can also pass a custom data frame to `ggcoef`. The following variables are expected:

- `term` (except if you customize the mapping)
- `estimate` (except if you customize the mapping)
- `conf.low` and `conf.high` (only if you want to display error bars)

```{r}
cust <- data.frame(
  term = c("male vs. female", "30-49 vs. 18-29", "50+ vs. 18-29", "urban vs. rural"),
  estimate = c(.456, 1.234, 1.897, 1.003),
  conf.low = c(.411, 1.042, 1.765, 0.678),
  conf.high = c(.498, 1.564, 2.034, 1.476),
  variable = c("sex", "age", "age", "residence")
)
cust$term <- factor(cust$term, cust$term)
ggcoef(cust, exponentiate = TRUE)
ggcoef(
  cust,
  exponentiate = TRUE,
  mapping = aes(x = estimate, y = term, colour = variable),
  size = 5
)
```
