---
title: "ggmatrix: plot matrix for ggplot2"
author: Barret Schloerke
date: Oct 29, 2015
output:
  html_document:
    highlight: default
    toc: true
    theme: united
vignette: >
  %\VignetteIndexEntry{ggmatrix}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r, echo=FALSE, message=FALSE, cache=FALSE}
knitr::read_chunk("_cache.R")
```
```{r cache, echo=FALSE, message=FALSE, cache=FALSE}
```
```{r, echo=FALSE, message=FALSE, cache=FALSE}
cache_wd("ggmatrix")
```

# ggmatrix

## Introduction

`ggmatrix` is a function for managing multiple plots in a matrix-like layout.  It was designed to adapt to any number of columns and rows.  This allows for very customized plot matrices.

## Generic Example

The examples below use plots labeled 1 to 6 to distinguish where the plots are being placed.  

```{r ggmatrix_genExample}
plotList <- list()
for (i in 1:6) {
  plotList[[i]] <- ggally_text(paste("Plot #", i, sep = ""))
}

# bare minimum of plotList, nrow, and ncol
pm <- ggmatrix(plotList, 2, 3)
pm

# provide more information
pm <- ggmatrix(
  plotList,
  nrow = 2, ncol = 3,
  xAxisLabels = c("A", "B", "C"),
  yAxisLabels = c("D", "E"),
  title = "Matrix Title"
)
pm

# display plots in column order
pm <- ggmatrix(
  plotList,
  nrow = 2, ncol = 3,
  xAxisLabels = c("A", "B", "C"),
  yAxisLabels = c("D", "E"),
  title = "Matrix Title",
  byrow = FALSE
)
pm
```

## Matrix Subsetting

Individual plots may be retrieved from the plot matrix and can be placed in the plot matrix.

```{r ggmatrix_place}
pm <- ggmatrix(
  plotList,
  nrow = 2, ncol = 3,
  xAxisLabels = c("A", "B", "C"),
  yAxisLabels = c("D", "E"),
  title = "Matrix Title"
)
pm
p2 <- pm[1,2]
p3 <- pm[1,3]
p2
p3
pm[1,2] <- p3
pm[1,3] <- p2
pm

```

## Themes

```{r ggmatrix_theme}
library(ggplot2)
pm <- ggmatrix(
  plotList,
  nrow = 2, ncol = 3,
  xAxisLabels = c("A", "B", "C"),
  yAxisLabels = c("D", "E"),
  title = "Matrix Title",
  byrow = FALSE
)
pm <- pm + theme_bw()
pm
```



## Axis Control

The X and Y axis have booleans to turn on/off the individual plot's axes on the bottom and left sides of the plot matrix.  To save time, `showAxisPlotLabels` can be set to override `showXAxisPlotLabels` and `showYAxisPlotLabels`.

```{r ggmatrix_axisControl}
pm <- ggmatrix(
  plotList, nrow = 2, ncol = 3,
  xAxisLabels = c("A", "B", "C"),
  yAxisLabels = c("D", "E"),
  title = "No Left Plot Axis",
  showYAxisPlotLabels = FALSE
)
pm
pm <- ggmatrix(
  plotList, nrow = 2, ncol = 3,
  xAxisLabels = c("A", "B", "C"),
  yAxisLabels = c("D", "E"),
  title = "No Bottom Plot Axis",
  showXAxisPlotLabels = FALSE
)
pm
pm <- ggmatrix(
  plotList, nrow = 2, ncol = 3,
  xAxisLabels = c("A", "B", "C"),
  yAxisLabels = c("D", "E"),
  title = "No Plot Axes",
  showAxisPlotLabels = FALSE
)
pm
```

## Strips Control

By default, the plots in the top row and the right most column will display top-side and right-side strips respectively (`showStrips = NULL`).  If all strips need to appear in each plot, `showStrips` may be set to `TRUE`. If all strips should not be displayed, `showStrips` may be set to `FALSE`.

```{r ggmatrix_stripControl}
data(tips, package = "reshape")
plotList <- list(
  qplot(total_bill, tip, data = subset(tips, smoker == "No" & sex == "Female")) +
    facet_grid(time ~ day),
  qplot(total_bill, tip, data = subset(tips, smoker == "Yes" & sex == "Female")) +
    facet_grid(time ~ day),
  qplot(total_bill, tip, data = subset(tips, smoker == "No" & sex == "Male")) +
    facet_grid(time ~ day),
  qplot(total_bill, tip, data = subset(tips, smoker == "Yes" & sex == "Male")) +
    facet_grid(time ~ day)
)


pm <- ggmatrix(
  plotList, nrow = 2, ncol = 2,
  yAxisLabels = c("Female", "Male"),
  xAxisLabels = c("Non Smoker", "Smoker"),
  title = "Total Bill vs Tip",
  showStrips = NULL # default
)
pm
pm <- ggmatrix(
  plotList, nrow = 2, ncol = 2,
  yAxisLabels = c("Female", "Male"),
  xAxisLabels = c("Non Smoker", "Smoker"),
  title = "Total Bill vs Tip",
  showStrips = TRUE
)
pm
pm <- ggmatrix(
  plotList, nrow = 2, ncol = 2,
  yAxisLabels = c("Female", "Male"),
  xAxisLabels = c("Non Smoker", "Smoker"),
  title = "Total Bill vs Tip",
  showStrips = FALSE
)
pm
```

## Custom Printing

Since ggmatrix is not a "true" ggplot2 object, spacing between the plot axis labels and the matrix labels is not perfect.  There are a few `print.ggmatrix` parameters to help with this: `leftWidthProportion`, `bottomHeightProportion`, and `spacingProportion`.  Each parameter refers to the proportional size of 'left plot axis label area', 'bottom plot axis label area', and 'spacing between plots' relative to a single plotting area.  So, if a single plot is 2 inches wide and `spacingProportion = 0.03`, then the horizontal space between each plot will be 0.06 inches.

```{r ggmatrix_customPrinting}
data(presidential, package = "ggplot2")
plotList <- list(
  qplot(name, data = presidential, geom = "bar", fill = party) + ylim(c(0,6)) + coord_flip(),
  qplot(party, data = presidential, geom = "bar", fill = party) + coord_flip()
)
pm <- ggmatrix(
  plotList,
  2, 1,
  yAxisLabels = c("Presidents", "Party"),
  xAxisLabels = c("Count"),
  title = "President Term Length"
)
pm # default spacing
# use print.ggmatrix parameters
print(pm, leftWidthProportion = 0.12, spacingProportion = 0.05)
```
