---
title: "Additional statistics for ggplot2"
output: rmarkdown::html_vignette
author: GGally team
date: May 28, 2020
vignette: >
  %\VignetteIndexEntry{Additional statistics for ggplot2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```


`GGally` proposes several additional statistics that could be used with `ggplot2`. As reminder, a statistic is always used in conjunction with a geometry. You can call a statistic from a `geom_*()` or call a geometry from a `stat_*()`. A statistic will compute new variables from the provided `data`. These new variables could be mapped to an aesthetic using `ggplot2::after_stat()`.


```{r}
library(GGally, quietly = TRUE)
```

## `stat_cross()`

This statistic is intended to be used with two discrete variables mapped to **x** and **y** aesthetics. It will compute several statistics of a cross-tabulated table using `broom::tidy.test()` and `stats::chisq.test()`. More precisely, the computed variables are:

- **observed**: number of observations in x,y
- **prop**: proportion of total
- **row.prop**: row proportion
- **col.prop**: column proportion
- **expected**: expected count under the null hypothesis
- **residuals**: Pearson's residual
- **stdres**: standardized residual

By default, `stat_cross()` is using `ggplot2::geom_points()`. If you can to plot the number of observations, you need to map `after_stat(observed)` to an aesthetic (here **size**):

```{r}
d <- as.data.frame(Titanic)
ggplot(d) +
  aes(x = Class, y = Survived, weight = Freq, size = after_stat(observed)) +
  stat_cross() +
  scale_size_area(max_size = 20)
```

Note that the **weight** aesthetic is taken into account by `stat_cross()`.

We can go further using a custom shape and filling points with standardized residual to identify visually cells who are over- or underrepresented.

```{r fig.height=6, fig.width=6}
ggplot(d) +
  aes(x = Class, y = Survived, weight = Freq, size = after_stat(observed), fill = after_stat(stdres)) +
  stat_cross(shape = 22) +
  scale_fill_steps2(breaks = c(-3, -2, 2, 3), show.limits = TRUE) +
  scale_size_area(max_size = 20)
```

We can easily recreate a cross-tabulated table.

```{r}
ggplot(d) +
  aes(x = Class, y = Survived, weight = Freq) +
  geom_tile(fill = "white", colour = "black") +
  geom_text(stat = "cross", mapping = aes(label = after_stat(observed))) +
  theme_minimal()
```

Even more complicated, we want to produce a table showing column proportions and where cells are filled with standardized residuals. Note that `stat_cross()` could be used with facets. In that case, computation is done separately in each facet.

```{r}
ggplot(d) +
  aes(
    x = Class, y = Survived, weight = Freq,
    label = scales::percent(after_stat(col.prop), accuracy = .1),
    fill = after_stat(stdres)
  ) +
  stat_cross(shape = 22, size = 30) +
  geom_text(stat = "cross") +
  scale_fill_steps2(breaks = c(-3, -2, 2, 3), show.limits = TRUE) +
  facet_grid(rows = vars(Sex)) +
  labs(fill = "Standardized residuals") +
  theme_minimal()
```


## `stat_weighted_mean()`

`stat_weighted_mean()` computes mean value of **y** (taking into account any **weight** aesthetic if provided) for each value of **x**. More precisely, it will return a new data frame with one line per unique value of **x** with the following new variables:

- **y**: mean value of the original **y** (i.e. **numerator**/**denominator**)
- **numerator**
- **denominator**

Let's take an example. The following plot shows all tips received accrding to the day of the week.

```{r}
data(tips, package = "reshape")
ggplot(tips) +
  aes(x = day, y = tip) +
  geom_point()
```

To plot their mean value per day, simply use `stat_weighted_mean()`.

```{r}
ggplot(tips) +
  aes(x = day, y = tip) +
  stat_weighted_mean()
```

We can specify the geometry we want using `geom` argument. Note that for lines, we need to specify the **group** aesthetic as well.

```{r}
ggplot(tips) +
  aes(x = day, y = tip, group = 1) +
  stat_weighted_mean(geom = "line")
```

An alternative is to specify the statistic in `ggplot2::geom_line()`.

```{r}
ggplot(tips) +
  aes(x = day, y = tip, group = 1) +
  geom_line(stat = "weighted_mean")
```

Of course, it could be use with other geometries. Here a bar plot.

```{r}
p <- ggplot(tips) +
  aes(x = day, y = tip, fill = sex) +
  stat_weighted_mean(geom = "bar", position = "dodge") +
  ylab("mean tip")
p
```

It is very easy to add facets. In that case, computation will be done separately for each facet.

```{r}
p + facet_grid(rows = vars(smoker))
```

`stat_weighted_mean()` could be also used for computing proportions as a proportion is technically a mean of binary values (0 or 1).

```{r}
ggplot(tips) +
  aes(x = day, y = as.integer(smoker == "Yes"), fill = sex) +
  stat_weighted_mean(geom = "bar", position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  ylab("proportion of smoker")
```

Finally, you can use the **weight** aesthetic to indicate weights to take into account for computing means / proportions.

```{r}
d <- as.data.frame(Titanic)
ggplot(d) +
  aes(x = Class, y = as.integer(Survived == "Yes"), weight = Freq, fill = Sex) +
  geom_bar(stat = "weighted_mean", position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Proportion who survived")
```

