---
title: "stat_*(): Additional statistics for ggplot2"
output: rmarkdown::html_vignette
author: GGally team
date: May 28, 2020
vignette: >
  %\VignetteIndexEntry{stat_*(): Additional statistics for ggplot2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 6,
  fig.height = 4
)
```


`GGally` proposes several additional statistics that could be used with `ggplot2`. As reminder, a statistic is always used in conjunction with a geometry. You can call a statistic from a `geom_*()` or call a geometry from a `stat_*()`. A statistic will compute new variables from the provided `data`. These new variables could be mapped to an aesthetic using `ggplot2::after_stat()`.


```{r}
library(GGally, quietly = TRUE)
```

## `stat_cross()`

This statistic is intended to be used with two discrete variables mapped to **x** and **y** aesthetics. It will compute several statistics of a cross-tabulated table using `broom::tidy.test()` and `stats::chisq.test()`. More precisely, the computed variables are:

- **observed**: number of observations in x,y
- **prop**: proportion of total
- **row.prop**: row proportion
- **col.prop**: column proportion
- **expected**: expected count under the null hypothesis
- **residuals**: Pearson's residual
- **stdres**: standardized residual

By default, `stat_cross()` is using `ggplot2::geom_points()`. If you can to plot the number of observations, you need to map `after_stat(observed)` to an aesthetic (here **size**):

```{r}
d <- as.data.frame(Titanic)
ggplot(d) +
  aes(x = Class, y = Survived, weight = Freq, size = after_stat(observed)) +
  stat_cross() +
  scale_size_area(max_size = 20)
```

Note that the **weight** aesthetic is taken into account by `stat_cross()`.

We can go further using a custom shape and filling points with standardized residual to identify visually cells who are over- or underrepresented.

```{r fig.height=6, fig.width=6}
ggplot(d) +
  aes(x = Class, y = Survived, weight = Freq, size = after_stat(observed), fill = after_stat(stdres)) +
  stat_cross(shape = 22) +
  scale_fill_steps2(breaks = c(-3, -2, 2, 3), show.limits = TRUE) +
  scale_size_area(max_size = 20)
```

We can easily recreate a cross-tabulated table.

```{r}
ggplot(d) +
  aes(x = Class, y = Survived, weight = Freq) +
  geom_tile(fill = "white", colour = "black") +
  geom_text(stat = "cross", mapping = aes(label = after_stat(observed))) +
  theme_minimal()
```

Even more complicated, we want to produce a table showing column proportions and where cells are filled with standardized residuals. Note that `stat_cross()` could be used with facets. In that case, computation is done separately in each facet.

```{r}
ggplot(d) +
  aes(
    x = Class, y = Survived, weight = Freq,
    label = scales::percent(after_stat(col.prop), accuracy = .1),
    fill = after_stat(stdres)
  ) +
  stat_cross(shape = 22, size = 30) +
  geom_text(stat = "cross") +
  scale_fill_steps2(breaks = c(-3, -2, 2, 3), show.limits = TRUE) +
  facet_grid(rows = vars(Sex)) +
  labs(fill = "Standardized residuals") +
  theme_minimal()
```

## `stat_prop()`

`stat_prop()` is a variation of `ggplot2::stat_count()` allowing to compute custom proportions according to the **by** aesthetic defining the denominator (i.e. all proportions for a same value of **by** will sum to 1). The **by** aesthetic should be a factor. Therefore, `stat_prop()` requires the **by** aesthetic and this **by** aesthetic should be a factor.

### adding labels on a percent stacked bar plot

When using `position = "fill"` with `geom_bar()`, you can produce a percent stacked bar plot. However, the proportions corresponding to the **y** axis are not directly accessible using only `ggplot2`. With `stat_prop()`, you can easily add them on the plot.

In the following example, we indicated `stat = "prop"` to `ggplot2::geom_text()` to use `stat_prop()`, we defined the **by** aesthetic (here we want to compute the proportions separately for each value of **x**), and we also used `ggplot2::position_fill()` when calling `ggplot2::geom_text()`.

```{r}
d <- as.data.frame(Titanic)
p <- ggplot(d) +
  aes(x = Class, fill = Survived, weight = Freq, by = Class) +
  geom_bar(position = "fill") +
  geom_text(stat = "prop", position = position_fill(.5))
p
```

Note that `stat_prop()` has properly taken into account the **weight** aesthetic.

`stat_prop()` is also compatible with faceting. In that case, proportions are computed separately in each facet.

```{r}
p + facet_grid(cols = vars(Sex))
```

### displaying proportions of the total

If you want to display proportions of the total, simply map the **by** aesthetic to `1`. Here an example using a stacked bar chart.

```{r}
ggplot(d) +
  aes(x = Class, fill = Survived, weight = Freq, by = 1) +
  geom_bar() +
  geom_text(
    aes(label = scales::percent(after_stat(prop), accuracy = 1)),
    stat = "prop",
    position = position_stack(.5)
 )
```

### a dodged bar plot to compare two distributions

A dodged bar plot could be used to compare two distributions.

```{r}
ggplot(d) +
  aes(x = Class, fill = Sex, weight = Freq, by = Sex) +
  geom_bar(position = "dodge")
```

On the previous graph, it is difficult to see if first class is over- or under-represented among women, due to the fact they were much more men on the boat. `stat_prop()` could be used to adjust the graph by displaying instead the proportion within each category (i.e. here the proportion by sex).

```{r}
ggplot(d) +
  aes(x = Class, fill = Sex, weight = Freq, by = Sex, y = after_stat(prop)) +
  geom_bar(stat = "prop", position = "dodge") +
  scale_y_continuous(labels = scales::percent)
```

The same example with labels:

```{r}
ggplot(d) +
  aes(x = Class, fill = Sex, weight = Freq, by = Sex, y = after_stat(prop)) +
  geom_bar(stat = "prop", position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  geom_text(
    mapping = aes(
      label = scales::percent(after_stat(prop), accuracy = .1), 
      y = after_stat(0.01)
    ),
    vjust = "bottom",
    position = position_dodge(.9),
    stat = "prop"
  )
```


## `stat_weighted_mean()`

`stat_weighted_mean()` computes mean value of **y** (taking into account any **weight** aesthetic if provided) for each value of **x**. More precisely, it will return a new data frame with one line per unique value of **x** with the following new variables:

- **y**: mean value of the original **y** (i.e. **numerator**/**denominator**)
- **numerator**
- **denominator**

Let's take an example. The following plot shows all tips received according to the day of the week.

```{r}
data(tips, package = "reshape")
ggplot(tips) +
  aes(x = day, y = tip) +
  geom_point()
```

To plot their mean value per day, simply use `stat_weighted_mean()`.

```{r}
ggplot(tips) +
  aes(x = day, y = tip) +
  stat_weighted_mean()
```

We can specify the geometry we want using `geom` argument. Note that for lines, we need to specify the **group** aesthetic as well.

```{r}
ggplot(tips) +
  aes(x = day, y = tip, group = 1) +
  stat_weighted_mean(geom = "line")
```

An alternative is to specify the statistic in `ggplot2::geom_line()`.

```{r}
ggplot(tips) +
  aes(x = day, y = tip, group = 1) +
  geom_line(stat = "weighted_mean")
```

Of course, it could be use with other geometries. Here a bar plot.

```{r}
p <- ggplot(tips) +
  aes(x = day, y = tip, fill = sex) +
  stat_weighted_mean(geom = "bar", position = "dodge") +
  ylab("mean tip")
p
```

It is very easy to add facets. In that case, computation will be done separately for each facet.

```{r}
p + facet_grid(rows = vars(smoker))
```

`stat_weighted_mean()` could be also used for computing proportions as a proportion is technically a mean of binary values (0 or 1).

```{r}
ggplot(tips) +
  aes(x = day, y = as.integer(smoker == "Yes"), fill = sex) +
  stat_weighted_mean(geom = "bar", position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  ylab("proportion of smoker")
```

Finally, you can use the **weight** aesthetic to indicate weights to take into account for computing means / proportions.

```{r}
d <- as.data.frame(Titanic)
ggplot(d) +
  aes(x = Class, y = as.integer(Survived == "Yes"), weight = Freq, fill = Sex) +
  geom_bar(stat = "weighted_mean", position = "dodge") +
  scale_y_continuous(labels = scales::percent) +
  labs(y = "Proportion who survived")
```

