---
title: "ggbivariate(): Plot an outcome with several potential explanatory variables"
output: rmarkdown::html_vignette
author: Joseph Larmarange
date: May 16, 2016
vignette: >
  %\VignetteIndexEntry{ggbivariate(): Plot an outcome with several potential explanatory variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GGally)
```


# `GGally::ggbivariate()`

The purpose of this function is to easily plot a visualization of the bivariate relation between one outcome and several explanatory variables.

## Basic example

Simply indicate the outcome and the explanatory variables. Both could be discrete or continuous.

```{r}
data(tips, package = "reshape")
ggbivariate(tips, outcome = "smoker", explanatory = c("day", "time", "sex", "tip"))
```

```{r}
ggbivariate(tips, outcome = "total_bill", explanatory = c("day", "time", "sex", "tip"))
```

If no explanatory variables are provided, will take all available variables other than the outcome.

```{r}
ggbivariate(tips, "smoker")
```


## Customize plot title and legend title

```{r}
ggbivariate(
  tips, "smoker", c("day", "time", "sex", "tip"),
  title = "Custom title"
) +
  labs(fill = "Smoker ?")
```

## Customize fill colour scale

```{r}
ggbivariate(tips, "smoker", c("day", "time", "sex", "tip")) +
  scale_fill_brewer(type = "qual")
```

## Customize labels

```{r}
ggbivariate(
  tips, "smoker", c("day", "time", "sex", "tip"),
  rowbar_args = list(
    colour = "white",
    size = 4,
    fontface = "bold",
    label_format = scales::label_percent(accurary = 1)
  )
)
```

## Choose the sub-plot from which to get the legend

```{r}
ggbivariate(tips, "smoker")
ggbivariate(tips, "smoker", legend = 3)
```

## Change theme

```{r}
ggbivariate(tips, "smoker") + theme_light()
```

## Use mapping to indicate weights

```{r}
d <- as.data.frame(Titanic)
ggbivariate(d, "Survived", mapping = aes(weight = Freq))
```

## Use types to customize types of subplots

```{r}
ggbivariate(
  tips, 
  outcome = "smoker", 
  explanatory = c("day", "time", "sex", "tip"),
  types = list(comboVertical = "autopoint")
)
```

For more customization options, you could directly use `ggduo()` (see also `vignette("ggduo")`).
