---
title: "ggcoef_model(): Model coefficients"
output: rmarkdown::html_vignette
author: Joseph Larmarange
date: May 16, 2016
vignette: >
  %\VignetteIndexEntry{ggcoef_model(): Model coefficients}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GGally)
```

# `GGally::ggcoef_model()`

The purpose of this function is to quickly plot the coefficients of a model. It is an updated and improved version of `GGally::ggcoef()` based on `broom.helpers::tidy_plus_plus()`. For displaying a nicely formatted table of the same models, look at `gtsummary::tbl_regression()`.

## Quick coefficients plot

To work automatically, this function requires the [`broom.helpers` package](https://larmarange.github.io/broom.helpers/). Simply call `ggcoef_model()` with a model object. It could be the result of `stats::lm`, `stats::glm` or any other model covered by `broom.helpers`[^1].

[^1]: See <https://larmarange.github.io/broom.helpers/articles/tidy.html>.

```{r ggcoef-reg}
data(tips, package = "reshape")
mod_simple <- lm(tip ~ day + time + total_bill, data = tips)
ggcoef_model(mod_simple)
```

In the case of a logistic regression (or any other model for which coefficients are usually exponentiated), simply indicated `exponentiate = TRUE`. Note that a logarithmic scale will be used for the x-axis.

```{r ggcoef-titanic}
d_titanic <- as.data.frame(Titanic)
d_titanic$Survived <- factor(d_titanic$Survived, c("No", "Yes"))
mod_titanic <- glm(
  Survived ~ Sex * Age + Class,
  weights = Freq,
  data = d_titanic,
  family = binomial
)
ggcoef_model(mod_titanic, exponentiate = TRUE)
```

## Customizing the plot

### Variable labels

You can use the `labelled` package to define variable labels. They will be automatically used by `ggcoef_model()`. Note that variable labels should be defined before computing the model.

```{r}
library(labelled)
tips_labelled <- tips %>%
  set_variable_labels(
    day = "Day of the week",
    time = "Lunch or Dinner",
    total_bill = "Bill's total"
  )
mod_labelled <- lm(tip ~ day + time + total_bill, data = tips_labelled)
ggcoef_model(mod_labelled)  
```

You can also define custom variable labels directly by passing a named vector to the `variable_labels` option.

```{r}
ggcoef_model(
  mod_simple,
  variable_labels = c(
    day = "Week day",
    time = "Time (lunch or dinner ?)",
    total_bill = "Total of the bill"
  )
)
```

If variable labels are to long, you can pass `ggplot2::label_crap_gen()` or any other labeller function to `facet_labeller.`

```{r}
ggcoef_model(
  mod_simple,
  variable_labels = c(
    day = "Week day",
    time = "Time (lunch or dinner ?)",
    total_bill = "Total of the bill"
  ),
  facet_labeller = label_wrap_gen(10)
)
```

Use `facet_row = NULL` to hide variable names.

```{r}
ggcoef_model(mod_simple, facet_row = NULL, colour_guide = TRUE)
```

When you have a model with several continuous variables (who have usually only one model term per variable), you can group variables displayed on one row together.

```{r}
mod <- lm(Sepal.Length ~ ., data = iris)
ggcoef_model(mod)
ggcoef_model(mod, group_one_row_variables = TRUE)
```

### Term labels

Several options allows you to customize term labels.

```{r}
ggcoef_model(mod_titanic, exponentiate = TRUE)
ggcoef_model(
    mod_titanic,
    exponentiate = TRUE,
    show_p_values = FALSE,
    signif_stars = FALSE,
    add_reference_rows = FALSE,
    categorical_terms_pattern = "{level} (ref: {reference_level})",
    interaction_sep = " x "
  ) +
  scale_y_discrete(labels = scales::label_wrap(15))
```

By default, for categorical variables using treatment and sum contrasts, reference rows will be added and displayed on the graph.

```{r}
mod_titanic2 <- glm(
  Survived ~ Sex * Age + Class,
  weights = Freq,
  data = d_titanic,
  family = binomial,
  contrasts = list(Sex = contr.sum, Class = contr.treatment(4, base = 3))
)
ggcoef_model(mod_titanic2, exponentiate = TRUE)
```

Use `no_reference_row` to indicate which variables should not have a reference row added.

```{r}
ggcoef_model(
  mod_titanic2, exponentiate = TRUE, 
  no_reference_row = "Sex"
)
ggcoef_model(
  mod_titanic2, exponentiate = TRUE, 
  no_reference_row = broom.helpers::all_dichotomous()
)
ggcoef_model(
  mod_titanic2, exponentiate = TRUE, 
  no_reference_row = broom.helpers::all_categorical(),
  categorical_terms_pattern = "{level}/{reference_level}"
)
```

### Elements to display

Use `intercept = TRUE` to display intercepts.

```{r}
ggcoef_model(mod_simple, intercept = TRUE)
```

You can remove confidence intervals with `conf.int = FALSE`.

```{r}
ggcoef_model(mod_simple, conf.int = FALSE)
```

By default, significant terms (i.e. with a p-value below 5%) are highlighted using two types of dots. You can control the level of significance with `significance` or remove it with `significance = NULL`.

```{r}
ggcoef_model(mod_simple, significance = NULL)
```

By default, dots are colored by variable. You can deactivate this behavior with `colour = NULL`.

```{r}
ggcoef_model(mod_simple, colour = NULL)
```

You can display only a subset of terms with **include**.

```{r}
ggcoef_model(mod_simple, include = c("time", "total_bill"))
```

It is possible to use `tidyselect` helpers.

```{r}
ggcoef_model(mod_simple, include = dplyr::starts_with("t"))
```

You can remove stripped rows with `stripped_rows = FALSE`.

```{r}
ggcoef_model(mod_simple, stripped_rows = FALSE)
```

Do not hesitate to consult the [help file](http://ggobi.github.io/ggally/reference/ggcoef_model.html) of `ggcoef_model()` to see all avalaible options.

## Multinomial models

For multinomial models, simply use `ggcoef_multinom()`. Two types of visualisation are available: "dodged" and "faceted".

```{r}
library(nnet)
data(happy)
mod <- multinom(happy ~ age + degree + sex, data = happy, trace = FALSE)
ggcoef_multinom(mod, exponentiate = TRUE)
ggcoef_multinom(mod, exponentiate = TRUE, type = "faceted")
ggcoef_multinom(
  mod, exponentiate = TRUE, 
  y.level = c(
      "pretty happy" = "pretty happy\n(ref: very happy)"
  )
)
```


## Comparing several models

You can easily compare several models with `ggcoef_compare()`.

```{r}
mod1 <- lm(Fertility ~ ., data = swiss)
mod2 <- step(mod1, trace = 0)
mod3 <- lm(Fertility ~ Agriculture + Education * Catholic, data = swiss)
models <- list("Full model" = mod1, "Simplified model" = mod2, "With interaction" = mod3)

ggcoef_compare(models)
ggcoef_compare(models, type = "faceted")
```


## Advanced users

Advanced users could use their own dataset and pass it to `ggcoef_plot()`. Such dataset could be produced by `ggcoef_model()`, `ggcoef_compare()` or `ggcoef_multinom()` with the option `return_data = TRUE` or by using `broom::tidy()` or `broom.helpers::tidy_plus_plus()`.
