---
title: "ggcoef(): Model coefficients"
output: rmarkdown::html_vignette
author: Joseph Larmarange
date: May 16, 2016
vignette: >
  %\VignetteIndexEntry{ggcoef(): Model coefficients}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GGally)
```

# `GGally::ggcoef()`

The purpose of this function is to quickly plot the coefficients of a model. For an updated and improved version, see `GGally::ggcoef_model()` and the [corresponding vignette](https://ggobi.github.io/ggally/articles/ggcoef_model.html).

## Quick coefficients plot

To work automatically, this function requires the `broom` package. Simply call `ggcoef()` with a model object. It could be the result of `stats::lm`, `stats::glm` or any other model covered by `broom` and its `broom::tidy` method[^1].

[^1]: See <http://www.rdocumentation.org/packages/broom>.

```{r ggcoef-reg}
reg <- lm(Sepal.Length ~ Sepal.Width + Petal.Length + Petal.Width, data = iris)
ggcoef(reg)
```

In the case of a logistic regression (or any other model for which coefficients are usually exponentiated), simply indicated `exponentiate = TRUE`. Note that a logarithmic scale will be used for the x-axis.

```{r ggcoef-titanic}
d <- as.data.frame(Titanic)
log.reg <- glm(Survived ~ Sex + Age + Class, family = binomial, data = d, weights = d$Freq)
ggcoef(log.reg, exponentiate = TRUE)
```

## Customizing the plot

You can use `conf.int`, `vline` and `exclude_intercept` to display or not confidence intervals as error bars, a vertical line for `x = 0` (or `x = 1` if coefficients are exponentiated) and the intercept.

```{r ggcoef-reg-custom}
ggcoef(reg, vline = FALSE, conf.int = FALSE, exclude_intercept = TRUE)
```

See the help page of `ggcoef()` for the full list of arguments that could be used to personalize how error bars and the vertical line are plotted.

```{r ggcoef-full-args}
ggcoef(
  log.reg,
  exponentiate = TRUE,
  vline_color = "red",
  vline_linetype =  "solid",
  errorbar_color = "blue",
  errorbar_height = .25
)
```

Additional parameters will be passed to [ggplot2::geom\_point()].

```{r ggcoef-log.reg}
ggcoef(log.reg, exponentiate = TRUE, color = "purple", size = 5, shape = 18)
```

Finally, you can also customize the aesthetic mapping of the points.

```{r ggcoef-aes}
library(ggplot2)
ggcoef(log.reg, exponentiate = TRUE, mapping = aes(x = estimate, y = term, size = p.value)) +
  scale_size_continuous(trans = "reverse")
```

## Custom data frame

You can also pass a custom data frame to [ggcoef()]. The following variables are expected:

-   `term` (except if you customize the mapping)
-   `estimate` (except if you customize the mapping)
-   `conf.low` and `conf.high` (only if you want to display error bars)

```{r ggcoef-data-frame}
cust <- data.frame(
  term = c("male vs. female", "30-49 vs. 18-29", "50+ vs. 18-29", "urban vs. rural"),
  estimate = c(.456, 1.234, 1.897, 1.003),
  conf.low = c(.411, 1.042, 1.765, 0.678),
  conf.high = c(.498, 1.564, 2.034, 1.476),
  variable = c("sex", "age", "age", "residence")
)
cust$term <- factor(cust$term, cust$term)
ggcoef(cust, exponentiate = TRUE)
ggcoef(
  cust,
  exponentiate = TRUE,
  mapping = aes(x = estimate, y = term, colour = variable),
  size = 5
)
```
