<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{ggnetworkmap}
-->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  error = FALSE,
  tidy = FALSE,
  fig.height = 7,
  fig.width = 9.3333,
  out.height = "600px",
  out.width = "800px"
)
```

# ggnetworkmap
#### Amos Elberg
#### January 10, 2015

This is the homepage for ``ggnetworkmap``, a function for plotting elegant maps using ``ggplot2``. ``ggnetworkmap`` is under consideration for inclusion in ``GGally``.

``ggnetworkmap`` builds on ``ggnet``, first by fixing the issue with ``ggnet`` that caused me to develop ``ggmapnetwork``: ``ggnet`` does not allow drawing the network over a map. ``ggnetworkmap`` does, and is particularly intended for use with ``ggmap``. Some examples appear below:

```{r echo = FALSE}
suppressMessages({require(ggplot2); require(igraph); require(maps); require(GGally)})
```

```{r}
require(ggplot2); require(igraph); require(maps); require(GGally)
set.seed(124)
airports <- read.csv("http://datasets.flowingdata.com/tuts/maparcs/airports.csv", header=TRUE)

airports$lon <- airports$long

flights <- data.frame(
  origin      = sample(airports[200:400,]$iata, 200, replace = TRUE),
  destination = sample(airports[200:400,]$iata, 200, replace = TRUE)
)

graph <- graph.data.frame(flights, airports, directed = TRUE)

# Subset the graph
graph <- graph - V(graph)[igraph::degree(graph, mode = "total") < 2]

# Add vertex attributes that will be used to color nodes
V(graph)$degree <- igraph::degree(graph, mode = "total")
V(graph)$mygroup <- sample(1:4, length(V(graph)), replace = TRUE)

# Create the map
usa <- ggplot(ggplot2::map_data("usa"), aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), color = "grey65", fill = "#f9f9f9", size = 0.2)

# Draw the plot
ggnetworkmap(
  usa,
  graph,
  size = 2,
  great.circles = TRUE,
  node.group = mygroup,
  ring.group = degree,
  segment.color = "cornflowerblue",
  weight = degree
)
```

This next example uses data from a spam community I identified on twitter while exploring and trying to clear-up a group of tweets. After coloring the nodes based on igraph's information centrality community detection function, the odd structure stood out clearly.

```{r}
data(twitter_spam_community)
world <- fortify(map("world", plot = FALSE, fill = TRUE))
gp <- ggplot(world, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), color = "grey65", fill = "#f9f9f9", size = 0.2)

ggnetworkmap(gp, twitter_spam_community)
```

Is the network really concentrated in the U.S.? Probably not. One of the odd things about the network, is a much higher proportion of the users gave locations that could be geocoded, than twitter users generally.

Let's see the network topology

```{r}
ggnetworkmap(data=twitter_spam_community)
```

That's an odd structure, let's see who's following who.

```{r}
ggnetworkmap(data=twitter_spam_community, arrow.size = 0.5)
```

Coloring nodes according to vertex attributes can highlight network structures.

```{r}
V(twitter_spam_community)$indegree <- igraph::degree(twitter_spam_community, mode = "in")
V(twitter_spam_community)$outdegree <- igraph::degree(twitter_spam_community, mode = "out")

ggnetworkmap(
  data = twitter_spam_community,
  arrow.size = 0.5,
  node.group = indegree,
  ring.group = outdegree, size = 4
) +
  scale_fill_continuous("Indegree", high = "red", low = "yellow") +
  labs(color = "Outdegree")
```

Some of the twitter data is already contained as vertex attributes

```{r}
ggnetworkmap(
  data = twitter_spam_community,
  arrow.size = 0.5,
  node.group = followersCount,
  ring.group = friendsCount,
  size = 4,
  weight = indegree,
  label.nodes = TRUE,
  vjust = -1.5
) +
  scale_fill_continuous("followersCount", high = "red", low = "yellow") +
  labs(color = "friendsCount") +
  scale_color_continuous(low = "lightgreen", high = "darkgreen")
```

