---
title: "ggnostic(): Model diagnostics plot matrix"
output: rmarkdown::html_vignette
author: Barret Schloerke
date: Oct 1, 2016
vignette: >
  %\VignetteIndexEntry{ggnostic(): Model diagnostics plot matrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GGally)
```

# `GGally::ggnostic()`


`ggnostic()` is a display wrapper to `ggduo()` that displays full model diagnostics for each given explanatory variable. By default, `ggduo()` displays the residuals, leave-one-out model sigma value, leverage points, and Cook's distance against each explanatory variable.  The rows of the plot matrix can be expanded to include fitted values, standard error of the fitted values, standardized residuals, and any of the response variables.  If the model is a linear model, stars are added according to the `stats::anova` significance of each explanatory variable.

Most diagnostic plots contain reference line(s) to help determine if the model is fitting properly

* **residuals:**
    * Type key: `".resid"`
    * A solid line is located at the expected value of 0 with dashed lines at the 95% confidence interval. ($0 \pm 1.96 * \sigma$)
    * Plot function: `ggally_nostic_resid()`.
    * See also `stats::residuals`
* **standardized residuals:**
    * Type key: `".std.resid"`
    * Same as residuals, except the standardized residuals equal the regular residuals divided by sigma.  The dashed lines are located at $0 \pm 1.96 * 1$.
    * Plot function: `ggally_nostic_std_resid()`.
    * See also `stats::rstandard`
* **leave-one-out model sigma:**
    * Type key: `".sigma"`
    * A solid line is located at the full model's sigma value.
    * Plot function: `ggally_nostic_sigma()`.
    * See also `stats::influence`'s value on `sigma`
* **leverage points:**
    * Type key: `".hat"`
    * The expected value for the diagonal of a hat matrix is $p / n$. Points are considered leverage points if they are large than $2 * p / n$, where the higher line is drawn.
    * Plot function: `ggally_nostic_hat()`.
    * See also `stats::influence`'s value on `hat`
* **Cook's distance:**
    * Type key: `".cooksd"`
    * Points that are larger than $4 / n$ line are considered highly influential points. Plot function: `ggally_nostic_cooksd()`. See also `stats::cooks.distance`
* **fitted points:**
    * Type key: `".fitted"`
    * No reference lines by default.
    * Default plot function: `ggally_points()`.
    * See also `stats::predict`
* **standard error of fitted points**:
    * Type key: `".se.fit"`
    * No reference lines by default.
    * Plot function: `ggally_nostic_se_fit()`.
    * See also `stats::fitted`
* **response variables:**
    * Type key: (response name in data.frame)
    * No reference lines by default.
    * Default plot function: `ggally_points()`.


## Life Expectancy Model Fitting

Looking at the dataset `datasets::state.x77`, we will fit a multiple regression model for Life Expectancy.

```{r life_model}
# make a data.frame and fix column names
state <- as.data.frame(state.x77)
colnames(state)[c(4, 6)] <- c("Life.Exp", "HS.Grad")
str(state)

# fit full model
model <- lm(Life.Exp ~ ., data = state)
# reduce to "best fit" model with
model <- step(model, trace = FALSE)
summary(model)
```

Next, we look at the variables for any high (|value| > 0.8) correlation values and general interaction behavior.
```{r nostic_scatmat}
# look at variables for high correlation (none)
ggscatmat(state, columns = c("Population", "Murder", "HS.Grad", "Frost"))
```

All variables appear to be ok.  Next, we look at the model diagnostics.
```{r nostic_diag}
# look at model diagnostics
ggnostic(model)
```

* The residuals appear to be normally distributed. There are a couple residual outliers, but 2.5 outliers are expected.
* There are 5 leverage points according the diagonal of the hat matrix
* There are 2 leverage points according to Cook's distance.  One is **much** larger than the other.

Let's remove the largest data point first to try and define a better model.

```{r nostic_no_hawaii}
# very high life expectancy
state[11, ]
state_no_hawaii <- state[-11, ]
model_no_hawaii <- lm(Life.Exp ~ Population + Murder + HS.Grad + Frost, data = state_no_hawaii)
ggnostic(model_no_hawaii)
```

There are no more outrageous Cook's distance values.  The model without Hawaii appears to be a good fitting model.

```{r nostic_summary}
summary(model)
summary(model_no_hawaii)
```

Since there is only a marginal improvement by removing Hawaii, the original model should be used to explain life expectancy.



## Full diagnostic plot matrix example

The following lines of code will display different modle diagnostic plot matrices for the same statistical model.  The first one is of the default settings. The second adds color according to the ``species``. Finally, the third displays all possible columns and uses `ggally_smooth()` to display the fitted points and response variables.

```{r nostic_flea}
flea_model <- step(lm(head ~ ., data = flea), trace = FALSE)
summary(flea_model)
# default output
ggnostic(flea_model)
# color'ed output
ggnostic(flea_model, mapping = ggplot2::aes(color = species))
# full color'ed output
ggnostic(
  flea_model,
  mapping = ggplot2::aes(color = species),
  columnsY = c("head", ".fitted", ".se.fit", ".resid", ".std.resid", ".hat", ".sigma", ".cooksd"),
  continuous = list(default = ggally_smooth, .fitted = ggally_smooth)
)
```
