---
title: "ggpairs(): Pairwise plot matrix"
output: rmarkdown::html_vignette
author: Barret Schloerke
date: Oct 29, 2015
vignette: >
  %\VignetteIndexEntry{ggpairs(): Pairwise plot matrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GGally)
```


# `GGally::ggpairs()`



`ggpairs()` is a special form of a `ggmatrix()` that produces a pairwise comparison of multivariate data. By default, `ggpairs()` provides two different comparisons of each pair of columns and displays either the density or count of the respective variable along the diagonal.  With different parameter settings, the diagonal can be replaced with the axis values and variable labels.

There are many hidden features within `ggpairs()`.  Please take a look at the examples below to get the most out of `ggpairs()`.

## Columns and Mapping

The `columns` displayed default to all columns of the provided `data`.  To subset to only a few columns, use the `columns` parameter.

```{r ggpairs_columns}
data(tips, package = "reshape")
pm <- ggpairs(tips)
pm
## too many plots for this example.

## reduce the columns being displayed
## these two lines of code produce the same plot matrix
pm <- ggpairs(tips, columns = c(1, 6, 2))
pm <- ggpairs(tips, columns = c("total_bill", "time", "tip"), columnLabels = c("Total Bill", "Time of Day", "Tip"))
pm
```

Aesthetics can be applied to every subplot with the `mapping` parameter.

```{r ggpairs_mapping}
library(ggplot2)
pm <- ggpairs(tips, mapping = aes(color = sex), columns = c("total_bill", "time", "tip"))
pm
```

Since the plots are default plots (or are helper functions from GGally), the aesthetic color is altered to be appropriate.  Looking at the example above, 'tip' vs 'total_bill' (pm[3,1]) needs the `color` aesthetic, while 'time' vs 'total_bill' needs the `fill` aesthetic.  If custom functions are supplied, no aesthetic alterations will be done.

## Matrix Sections

There are three major sections of the pairwise matrix: `lower`, `upper`, and `diag`.  The `lower` and `upper` may contain three plot types: `continuous`, `combo`, and `discrete`.  The 'diag' only contains either `continuous` or `discrete`.

* `continuous`: both X and Y are continuous variables
* `combo`: one X and Y variable is discrete while the other is continuous
* `discrete`: both X and Y are discrete variables

To make adjustments to each section, a list of information may be supplied.  The list can be comprised of the following elements:

* `continuous`:
    * a character string representing the tail end of a `ggally_NAME` function, or a custom function
    * current valid `upper$continuous` and `lower$continuous` character strings: `'points'`, `'smooth'`, `'density'`, `'cor'`, `'blank'`
    * current valid `diag$continuous` character strings: `'densityDiag'`, `'barDiag'`, `'blankDiag'`
* `combo`:
    * a character string representing the tail end of a `ggally_NAME` function, or a custom function. (not applicable for a `diag` list)
    * current valid `upper$combo` and `lower$combo` character strings: `'box'`, `'dot'`, `'facethist'`, `'facetdensity'`, `'denstrip'`, `'blank'`
* `discrete`:
    * a character string representing the tail end of a `ggally_NAME` function, or a custom function
    * current valid `upper$discrete` and `lower$discrete` character strings: `'ratio'`, `'facetbar'`, `'blank'`
    * current valid `diag$discrete` character strings: `'barDiag'`, `'blankDiag'`
* `mapping`:  if mapping is provided, only the section's mapping will be overwritten

```{r ggpairs_section}
library(ggplot2)
pm <- ggpairs(
  tips, columns = c("total_bill", "time", "tip"),
  lower = list(
    continuous = "smooth",
    combo = "facetdensity",
    mapping = aes(color = time)
  )
)
pm
```

A section list may be set to the character string `"blank"` or `NULL` if the section should be skipped when printed.

```{r ggpairs_blank}
pm <- ggpairs(
  tips, columns = c("total_bill", "time", "tip"),
  upper = "blank",
  diag = NULL
)
pm
```

## Custom Functions

The `ggally_NAME` functions do not provide all graphical options.  Instead of supplying a character string to a `continuous`, `combo`, or `discrete` element within `upper`, `lower`, or `diag`, a custom function may be given.

The custom function should follow the api of

```{r ggally_custom_function}
custom_function <- function(data, mapping, ...){
  # produce ggplot2 object here
}
```

There is no requirement to what happens within the function, as long as a ggplot2 object is returned.

```{r ggpairs_custom_function}
my_bin <- function(data, mapping, ..., low = "#132B43", high = "#56B1F7") {
  ggplot(data = data, mapping = mapping) +
    geom_bin2d(...) +
    scale_fill_gradient(low = low, high = high)
}
pm <- ggpairs(
  tips, columns = c("total_bill", "time", "tip"),
  lower = list(
    continuous = my_bin
  )
)
pm
```

## Function Wrapping

The examples above use default parameters to each of the subplots.  One of the immediate parameters to be set it `binwidth`.  This parameters is only needed in the lower, combination plots where one variable is continuous while the other variable is discrete.

To change the default parameter `binwidth` setting, we will `wrap()` the function.  `wrap()` first parameter should be a character string or a custom function.  The remaining parameters supplied to wrap will be supplied to the function at run time.

```{r ggpairs_wrap}
pm <- ggpairs(
  tips, columns = c("total_bill", "time", "tip"),
  lower = list(
    combo = wrap("facethist", binwidth = 1),
    continuous = wrap(my_bin, binwidth = c(5, 0.5), high = "red")
  )
)
pm
```

To get finer control over parameters, please look into custom functions.


## Plot Matrix Subsetting

Please look at the [vignette for ggmatrix](#ggallyggmatrix) on plot matrix manipulations.

Small `ggpairs()` example:

```{r ggpairs_matrix}
pm <- ggpairs(tips, columns = c("total_bill", "time", "tip"))
# retrieve the third row, first column plot
p <- pm[3,1]
p <- p + aes(color = time)
p
pm[3,1] <- p
pm
```

## Themes

Please look at the [vignette for ggmatrix](#ggallyggmatrix) on plot matrix manipulations.

Small `ggpairs()` example:

```{r ggpairs_theme}
pmBW <- pm + theme_bw()
pmBW
```

## References

John W Emerson, Walton A Green, Barret Schloerke, Jason Crowley, Dianne Cook, Heike Hofmann, Hadley Wickham.
**[The Generalized Pairs Plot](http://vita.had.co.nz/papers/gpp.html)**.
*Journal of Computational and Graphical Statistics*, vol. 22, no. 1, pp. 79-91, 2012.
