---
title: "ggsurv(): Survival curves"
output: rmarkdown::html_vignette
author: Edwin Thoen
date: April, 4, 2016
vignette: >
  %\VignetteIndexEntry{ggsurv(): Survival curves}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(GGally)
```

# `GGally::ggsurv()`



This function produces Kaplan-Meier plots using `ggplot2`.
As a first argument, `ggsurv()` needs a `survival::survfit()` object.
Default settings differ for single stratum and multiple strata objects.

## Single Stratum
```{r basic-usage, fig.height=7, fig.width=7}
require(ggplot2)
require(survival)
require(scales)

data(lung, package = "survival")
sf.lung <- survival::survfit(Surv(time, status) ~ 1, data = lung)
ggsurv(sf.lung)
```

## Multiple Stratum

The legend color positions matches the survival order or each stratum, where the stratums that end at a lower value or time have a position that is lower in the legend.

```{r ggsurv-multiple}
sf.sex <- survival::survfit(Surv(time, status) ~ sex, data = lung)
pl.sex <- ggsurv(sf.sex)
pl.sex
```

## Alterations

Since a ggplot2 object is returned, plot objects may be altered after the original creation.

### Adjusting the legend

```{r ggsurv-legend}
pl.sex +
  ggplot2::guides(linetype = "none") +
  ggplot2::scale_colour_discrete(
    name   = "Sex",
    breaks = c(1, 2),
    labels = c("Male", "Female")
  )
```

### Adjust the limits

```{r ggsurv-limits}
data(kidney, package = "survival")
sf.kid <- survival::survfit(Surv(time, status) ~ disease, data = kidney)
pl.kid <- ggsurv(sf.kid, plot.cens = FALSE)
pl.kid

# Zoom in to first 80 days
pl.kid + ggplot2::coord_cartesian(xlim = c(0, 80), ylim = c(0.45, 1))
```

### Add text and remove the legend
```{r ggsurv-text}
pl.kid +
  ggplot2::annotate(
    "text",
    label = c("PKD", "Other", "GN", "AN"),
    x = c(90, 125, 5, 60),
    y = c(0.8, 0.65, 0.55, 0.30),
    size = 5,
    colour = scales::hue_pal(
      h         = c(0, 360) + 15,
      c         = 100,
      l         = 65,
      h.start   = 0,
      direction = 1
    )(4)
  ) +
  ggplot2::guides(color = "none", linetype = "none")
```
